---
title: "Covid Data Preparation & Processing"
output: html_notebook
---

```{r}
# Download libraries
library(tidyverse)
library(skimr)

covid_data <- read.csv("C:/Users/Snow/Documents/Data analytics/covid19_deaths_analysis/owid-covid-data.csv")

head(covid_data)
```

```{r}
# Summary of the data

skim_without_charts(covid_data)
```

We have 238 country registered while there should be in fact 195 country, this proves that there are some errors in the data, after further exploration I noticed that some
Also, the date is a character while it should be a date, so we will start by converting it to a date.

```{r}

covid_data<-mutate(covid_data,date = as.Date(date))

```

## 1. Covid deaths

Next we are creating our first tibble which will include covid deaths data:

```{r}
library(dplyr)

covid_deaths_tibble <- tibble(iso_code=covid_data$iso_code, continent=covid_data$continent, location=covid_data$location, date=covid_data$date,new_deaths= covid_data$new_deaths, total_deaths=covid_data$total_deaths, population=covid_data$population)

skim_without_charts(covid_deaths_tibble)

```

Since it some values are NA in total_deaths and new_deaths, we will replace them by 0, however we will delete the rows where the population is NA.

```{r}

covid_deaths_tibble <- covid_deaths_tibble %>%
                      mutate(new_deaths=if_else(is.na(new_deaths),0, new_deaths),
                      total_deaths=if_else(is.na(total_deaths),0, total_deaths))

covid_deaths_tibble <- drop_na(covid_deaths_tibble)
skim_without_charts(covid_deaths_tibble)
```

Now we can save the first csv file

```{r}
write_csv(covid_deaths_tibble,"covid_deaths_jan2020_Feb2022.csv")

```

## 2. Covid Vaccines 
Next, we will treat the vaccinations

```{r}

covid_vaccinations_tibble <- tibble(iso_code=covid_data$iso_code, continent=covid_data$continent, location=covid_data$location, date=covid_data$date,new_vaccinations= covid_data$new_vaccinations, total_vaccinations=covid_data$total_vaccinations,people_vaccinated=covid_data$people_vaccinated, people_fully_vaccinated= covid_data$people_fully_vaccinated, population=covid_data$population)

head(covid_vaccinations_tibble)
```

Similarly to what we did with the first tibble, we will remove the NA values and replace them with 0 for most columns, however we will delete the rows where the population is NA.

```{r}

covid_vaccinations_tibble <- covid_vaccinations_tibble %>%
                      mutate(new_vaccinations=if_else(is.na(new_vaccinations),0, new_vaccinations), total_vaccinations=if_else(is.na(total_vaccinations),0, total_vaccinations), people_vaccinated=if_else(is.na(people_vaccinated),0, people_vaccinated), people_fully_vaccinated=if_else(is.na(people_fully_vaccinated),0, people_fully_vaccinated))

covid_vaccinations_tibble <- drop_na(covid_vaccinations_tibble)

skim_without_charts(covid_vaccinations_tibble)

```

Now we can save the csv file

```{r}
write_csv(covid_vaccinations_tibble,"covid_vaccinations_jan2020_Feb2022.csv")

```

## 3. Covid Tests
Next data is the covid tests

```{r}

covid_tests_tibble <- tibble(iso_code=covid_data$iso_code, continent=covid_data$continent, location=covid_data$location, date=covid_data$date,new_tests=covid_data$new_tests, total_tests=covid_data$total_tests,positive_rate=covid_data$positive_rate, population=covid_data$population)

head(covid_tests_tibble)

```

We can see that there is a lot of NA values, we will start by replacing them with 0's and delete NA values of population

```{r}

covid_tests_tibble <- covid_tests_tibble %>%
                      mutate(new_tests=if_else(is.na(new_tests),0, new_tests), total_tests=if_else(is.na(total_tests),0, total_tests), positive_rate=if_else(is.na(positive_rate),0, positive_rate))

covid_tests_tibble <- drop_na(covid_tests_tibble)

skim_without_charts(covid_tests_tibble)

```


Now we can save the csv file

```{r}
write_csv(covid_tests_tibble,"covid_tests_jan2020_Feb2022.csv")

```

## 4. Country information
Here I want some information about the age, health conditions of each population
```{r}

countries_demographics_tibble <- tibble(iso_code=covid_data$iso_code, continent=covid_data$continent, date= covid_data$date,location=covid_data$location, population=covid_data$population,stringency_index=covid_data$stringency_index,
population_density=covid_data$population_density,reproduction_rate=covid_data$reproduction_rate, median_age=covid_data$median_age,aged_65_older=covid_data$aged_65_older, aged_70_older=covid_data$aged_70_older,extreme_poverty=covid_data$extreme_poverty, cardiovasc_death_rate=covid_data$cardiovasc_death_rate, diabetes_prevalence=covid_data$diabetes_prevalence, female_smokers=covid_data$female_smokers, male_smokers=covid_data$male_smokers, life_expectancy=covid_data$life_expectancy)

skim_without_charts(countries_demographics_tibble)

```

For this data I will just drop all NA and remove duplicates

```{r}
countries_demographics_tibble <- drop_na(countries_demographics_tibble)
countries_demographics_tibble <- unique(countries_demographics_tibble)
skim_without_charts(countries_demographics_tibble)
```

Lastly we will save this data as a new csv

```{r}
write_csv(countries_demographics_tibble,"countries_demographics_jan2020_Feb2022.csv")

```